Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-02-27 manuscript/HIAs.qmd
minor clean

---
title: Using spatially biased variables in ecosystem condition accounting with a GIS based workflow
subtitle: A Short Subtitle
author:
  - name: Anders Lorentzen Kolstad
    email: anders.kolstad@nina.no
    affiliations: 
        - id: some-tech
          name: Norwegian Institute for Nature Research
          department: Department of Terrestrial Ecology
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
    note: Yet another author footnote.
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - SEEA EA
  - alien species
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Ecological Indicators
      formatting: preprint
      model: 3p
      cite-style: authoryear
    include-in-header:
      text: '\usepackage{lineno}\linenumbers'
editor: 
  markdown: 
    wrap: sentence
execute:
  echo: false
  include: false
  eval: false
  warning: false
  message: false
---

<!--# The document is set to eval=F in the yaml. This only affect the rendering, and not the interactive session. Eval is also set explicitly for each chunk. When eval is set to false it means the job is quite time consuming and I instead run it once and cache it locally.  -->

@@

library(tidyverse)
library(knitr)
library(sf)
library(tmap)
library(tmaptools)
library(stars)
library(terra)
library(tidyterra)
library(ggtext)
library(cowplot)
library(units)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggmagnify)

@@

myVars <- c("7TK", "7SE", "PRTK", "PRSL", "7FA", "7GR-GI")
nts <- naturetypes_summary %>%
  rowwise() %>%
  mutate(keepers = sum(c_across(
    all_of(myVars))>0, na.rm=T)) |>
  filter(keepers >0,
         Ecosystem == "våtmark") |>
  pull(Nature_type)
```

@@

# Now I will copy these ADVS-values into the sf object again, keeping things in wide format
naturetypes <- naturetypes |>
  pivot_wider(
    names_from = "variable",
    values_from = "value"
  ) |>
  left_join(naturetypes_wide |> select(id, ADSV), by = "id") |>
  select(!c("7TK", "7SE", "PRSL", "PRTK"))
  # >
  #pivot_longer(cols = c(ADSV, "7FA", "7GR-GI"),
  #             names_to = "variable",
  #             values_drop_na = T)
head(naturetypes)
```

@@

# For 7GR-GI I use this
upper2 <- 1
lower2 <- 5
threshold2 <- 2.5 # = observable effect. Value 3 indicates a shit to a new type (grunntype)

@@

```{r}
#| eval: true
# Adding scaled indicator values to the dataset
naturetypes$i_ADSV <- eaTools::ea_normalise(
  data = naturetypes,
  vector = "ADSV",
  upper_reference_level = lower,
  lower_reference_level = upper,
  break_point = threshold,
  reverse = T
)

@@

# The data contains some multisurfaces as well
# table(st_geometry_type(muni))
# function to make sure that multipolygons are returned
ensure_multipolygons <- function(X) {
  tmp1 <- tempfile(fileext = ".gpkg")
  tmp2 <- tempfile(fileext = ".gpkg")
  st_write(X, tmp1)
  gdalUtilities::ogr2ogr(tmp1, tmp2, f = "GPKG", nlt = "MULTIPOLYGON")
  Y <- st_read(tmp2)
  st_sf(st_drop_geometry(X), geom = st_geometry(Y))
}

@@

# and also to get the data coverage polygon.
coverage3 <- coverage |>
  st_intersection(muni3)

## These I had to manually download from naturbase.no
#dk <- read_sf("../data/dekningskartNF/Naturtyper NiN - Dekningskart.shp") |>
#  rbind(read_sf("../data/dekningskartTR/Naturtyper NiN - Dekningskart.shp")) |>
#  rbind(read_sf("../data/dekningskartME/Naturtyper NiN - Dekningskart.shp")) |>
#  st_transform(st_crs(muni3)) |>
#  st_intersection(muni3) |>
#  st_union()
```

@@

```{r centroids}
#| eval: true
# Make map to show where the three municipalities are
# First, get centroids
centroids <- muni3 |>
  st_centroid()
```

```{r positionMap}
#| eval: true

world <- ne_countries(scale = "medium", returnclass = "sf") |>
  st_transform(myCRS) |>
  filter(admin %in% c("Norway", "Sweden")) |>
  st_make_valid()

@@

# what is the distance between Nordre Follo and Nord-Aurdal
(km_distance <- centroids |>
  st_distance() |>
  max() |>
  set_units("km") |>
  drop_units() |>
  round())
```

@@

# I first tried to import and crop the mire data using stars, but that failed (see pre 21 feb 2023)
# trying { terra } instead

@@

# Plot to check overlap
# ggplot()+
#  geom_spatraster(data = mire_nf_terra)+
#  geom_spatvector(data = nf_vect,
#                  fill = NA)
# The cropping and masking worked.
# 
# # I like the stars, sf and tmap combo better, so I return to stars
mire_stars_nf <- mire_terra_nf |>
  st_as_stars()
mire_stars_gr <- mire_terra_gr |>
  st_as_stars()
mire_stars_na <- mire_terra_na |>
  st_as_stars()

@@

# Make a plot to check that it has worked
infraMuni3_tbl |>
  ggplot() +
  geom_bar(
    aes(
      x = infrastructureIndex,
      y = total_area_HIAs_km2,
      fill = factor(infrastructureIndex),
      colour = factor(infrastructureIndex)
    ),
    stat = "identity",
    lwd = 1.2
  ) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(4, "YlOrBr")) +
  scale_color_manual(values = RColorBrewer::brewer.pal(5, "YlOrBr")[-1]) +
  theme_minimal_hgrid() +
  labs(
    x = "Human Impact Area (HIA) class",
    y = "Area (km<sup>2</sup>)"
  ) +
  theme(
    axis.title.x = element_textbox_simple(
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
      r = grid::unit(8, "pt"),
      fill = "azure1"
    ),
    axis.title.y = element_textbox_simple(
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
      orientation = "left-rotated",
      r = grid::unit(8, "pt"),
      fill = "azure1"
    ),
    strip.background = element_blank(),
    strip.text = element_textbox(
      size = 12,
      color = "white", fill = "#5D729D", box.color = "#4A618C",
      halign = 0.5, linetype = 1, r = unit(5, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3)
    )
  ) +
  guides(fill = "none", colour = "none") +
  scale_y_log10() +
  facet_grid(cols = vars(Municipality))
```

```{r corrCheck}
#| eval: true
#| cache: true
# Now I want to see if the indicator values depend on the HIA is a predictable
# way to justify the stratification
corrCheck <- st_intersection(naturetypes, infra)

# corrCheck |>
#  as_tibble() |>
#  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
#         i_alien_fct = floor(round(i_alien*10,2))/10) |>
#  count(i_ADSV_fct, infrastructureIndex) |>
#  View()

```{r HIA-validate}
#| eval: false
# A first look
corrCheck |>
  mutate(
    i_ADSV_fct = floor(round(i_ADSV * 10, 2)) / 10,
    i_alien_fct = floor(round(i_alien * 10, 2)) / 10,
    i_ditch_fct = floor(round(i_ditch * 10, 2)) / 10
  ) |>
  pivot_longer(
    cols = c(i_ADSV_fct, i_alien_fct, i_ditch_fct),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = factor(indicatorValue)
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_bw(base_size = 12) +
  guides(fill = guide_legend("Scaled indicator values")) +
  ylab("Fraction of data points") +
  xlab("HIA") +
  scale_fill_brewer(palette = "RdYlGn") +
  facet_grid(indicator ~ year)
```

```{r test 4}
#| eval: false
# The year effect is constant, so ignoring that
corrCheck |>
  mutate(
    i_ADSV_fct = floor(round(i_ADSV * 10, 2)) / 10,
    i_alien_fct = floor(round(i_alien * 10, 2)) / 10,
    i_ditch_fct = floor(round(i_ditch * 10, 2)) / 10
  ) |>
  pivot_longer(
    cols = c(i_ADSV_fct, i_alien_fct, i_ditch_fct),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = factor(indicatorValue)
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_bw(base_size = 12) +
  guides(fill = guide_legend("Scaled indicator values")) +
  ylab("Fraction of data points") +
  xlab("HIA") +
  scale_fill_brewer(
    palette = "RdYlGn") +
  facet_grid(~indicator)
```

@@

```{r test3}
#| eval: false
# The figure changes a lot depending on which value gets the darker colour
corrCheck |>
  pivot_longer(
    cols = c(i_ADSV, i_alien, i_ditch),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  mutate(
    condition = case_when(
      indicatorValue < 0.6 ~ "Poor",
      indicatorValue < 0.8 ~ "Intermediate",
      indicatorValue < 0.91 ~ "Good",
      .default = "Perfect"
    ),
    condition = fct_reorder(condition, -indicatorValue)
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = condition
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_bw(base_size = 12) +
  guides(fill = guide_legend("Scaled indicator values")) +
  ylab("Fraction of data points") +
  xlab("HIA") +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)
```

```{r test2}
#| eval: false
# If I look at this on a log scale, it also changes and becomes less convincing for ADSV and alien
corrCheck |>
  pivot_longer(
    cols = c(i_ADSV, i_alien, i_ditch),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  mutate(
    condition = case_when(
      indicatorValue < 0.6 ~ "Poor",
      indicatorValue < 0.8 ~ "Intermediate",
      indicatorValue < 0.91 ~ "Good",
      .default = "Perfect"
    ),
    condition = fct_reorder(condition, -indicatorValue)
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = condition
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_bw(base_size = 12) +
  guides(fill = guide_legend("Scaled indicator values")) +
  ylab("Fraction of data points") +
  xlab("HIA") +
  scale_y_log10() +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)

# This is not correct I think as it evens out differences that are clearly there.
# i_ditch is shown less of a pattern for example
```

```{r test}
#| eval: false

# HIAs 1 and 2 are quite similar for i_alien and i_ADSV, so perhaps I can combine them
corrCheck |>
  pivot_longer(cols = c(i_ADSV, i_alien),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  mutate(condition = case_when(
    indicatorValue <0.6 ~"Poor",
    indicatorValue <0.8 ~"Intermediate",
    indicatorValue <0.91 ~"Good",
    .default = "Perfect"
  ),
  condition = fct_reorder(condition, indicatorValue),
  HIA = case_when(
    infrastructureIndex %in% c(1,2) ~ "1 & 2",
    .default = factor(infrastructureIndex)
  )
  ) |>
  ggplot(aes(
    x = HIA, 
    fill = condition))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_y_log10() +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)
```

@@

#tmap_save(methodsMap, "../images/studyLocations.tiff",
#           dpi= 1000,
#           units = "cm",
#           width = 18,
#           height = 10)
#tmap_save(methodsMap, "../images/studyLocations.jpg",
#           units = "cm",
#           width = 18,
#           height = 10)

#
#saveRDS(methodsMap, "../figures/studyLocation.RDS")
```

@@

# Number of mire polygons per municipality
# -- Try count() here instead
nature3_tbl <- nature3 |>
  group_by(Municipality) |>
  summarise(n = n()) |>
  as_tibble()

# Take muni3 and add all sorts of other data to it
# using left_join.
# Save local cached file
muni_tbl <- muni3 |>
  # calculate total area
  mutate(
    area_km =
      round(
        units::drop_units(
          geom |> st_area() * 1e-6
        )
      )
  ) |>
  # make tibble for the coming join
  as_tibble() |>
  # paste inn terrestrial area
  left_join(
    terrestrial |>
      as_tibble() |>
      select(kommunenummer, t_area_km),
    keep = F
  ) |>
  # area of survey
  left_join(dk2 |> select(Municipality, dk_area_km)) |>
  mutate(dk_percent = round((dk_area_km / t_area_km) * 100)) |>
  # number of polygons
  left_join(nature3_tbl |> select(Municipality, n)) |>
  # total mire area and %
  left_join(mireArea |> select(Municipality, mirePercent, mire_km2)) |>
  # % mire inside survey coverage map
  left_join(mire_in_dk |> select(Municipality, mireInSurvey_km2)) |>
  mutate(mireInSurvery_percent = round(mireInSurvey_km2 / mire_km2 * 100, 2)) |>
  left_join(infraMuni3_summary)
```

```{r read-munitbl}
#| eval: true
#muni_tbl <- readRDS("../data/municipality_stats.RDS")
```

@@

```{r forets plot}
# Forest plot
# https://stackoverflow.com/questions/58657802/forest-plot-with-subgroups-in-ggplot2
# 
# #define colours for dots and bars
dotCOLS = c("grey90","grey70", "grey50", "grey40")
barCOLS <- tmaptools::get_brewer_pal("Set1", n = 4)

@@

```{r}
# Next I need to 
#  - copy the weighted means and SDs over to  HIA map and then over to the 
#    ecosystem delineation map polygons. 
#  - show intermittent resulting map with ggmagnify
#  - sample from these distributions (ignore missing HIAs)
```

@@

spread_na_map <- spread_na |>
  filter(indicator == "i_ditch") |>
  mutate(Trenching = factor(round(mean, 2))) |>
  ggplot() +
  geom_sf(aes(fill=Trenching,
              color = Trenching)) +
  geom_sf(data = na,
          alpha=0) +
  scale_fill_manual(values = myCols) +
  scale_color_manual(values = myCols) +
  coord_sf(
    datum = st_crs(myCRS),
    xlim = c(494174.8 , 537114.7 ), 
    ylim = c(6737092 , 6789676)) +
  ggmagnify::geom_magnify(from = from, to = to, 
                          expand = 0,
                          shadow =T,
                          corners = 0.1) +
  theme_bw()

#ggsave("../images/spread-na.tiff",
#       plot = spread_na_map)
#ggsave("../images/spread-na-small.tiff",
#       plot = spread_na_map,
#       dpi=150)
#ggsave("../images/spread-na-small.jpg",
#       plot = spread_na_map)
```

@@

# Now I sample indicator values from the EDM and create a new distribution for 
# indicator values in the EEA (the municipalities). First I sample the individual
# distributions for each mire polygon with n defined by the polygon area.
# The distribution for the polygon areas is strongly right skewed, meaning
# some polygons will contribute much more to the EAA value than others.
# I will keep this design in-line with SEEA EA guidelines for area weighting, 
# but it's worth noting that this could be solved in other ways.
# 
combineAll <- rbind(
  spread_nf,
  spread_gr,
  spread_na
) |>
  as_tibble() |>
  drop_na() |>
  group_by(Municipality, indicator) |>
  rowwise() |>
  mutate(
    i_sample = list(sample(rnorm(n, mean, sd)))) |>
  select(-geometry) |>
  group_by(Municipality, indicator) |>
  reframe(i_sample = 
            sample(
              unlist(i_sample), 
              size = 1000,
              replace = TRUE)) |>
  mutate(i_sample = case_when(
    i_sample < 0 ~ 0,
    i_sample >1  ~ 1,
    .default = i_sample
  ))
```

```{r EEA-plot}
#| eval: true
#| cache: true
library(ggridges)

(ridgepPlot <- combineAll |>
  mutate(
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )) |>
  ggplot(
    aes(
      x = i_sample, 
      y = Municipality,
      fill = after_stat(x))
  ) +
  geom_density_ridges_gradient(
    bandwidth=.05,
    quantile_lines = TRUE, 
    quantiles = c(0.025, .5, 0.975)
  ) +
  #scale_fill_viridis_c(option = "D") +
  scale_fill_distiller(palette = "RdYlGn",
    direction = 1) +
  theme_bw(base_size = 15) +
  guides(fill = "none") +
  geom_vline(xintercept = .6,
    size=1.2,
    lty=2) +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
  labs(y = "", x = "Indicator values")+
  xlim(c(0,1)) +
  facet_wrap(.~indicator)
)

@@

# saveRDS(EEA_tbl, "../output/EEA-table.RDS")
```

>>>>

---
title: Using spatially biased variables in ecosystem condition accounting with a GIS based workflow
subtitle: A Short Subtitle
author:
  - name: Anders Lorentzen Kolstad
    email: anders.kolstad@nina.no
    orchid: https://orcid.org/0000-0002-9623-9491
    affiliations: 
        - id: nina
          name: Norwegian Institute for Nature Research
          department: Department of Terrestrial Ecology
          address: Pb 5685 Torgarden
          city: Trondheim
          postal-code: 7485
    attributes:
        corresponding: true
  - name: Matthew Grainger
    email: matthew.grainger@nina.no
    orchid: https://orcid.org/0000-0001-8426-6495
    affiliations:
        - ref: nina
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - alien species
  - disturbance
  - indicator
  - ecosystem condition
  - wetlands
  - mire
  - peatlands
  - ecosystem accounting
  - SEEA EA
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Ecological Indicators
      formatting: preprint
      model: 3p
      cite-style: authoryear
    include-in-header:
      text: '\usepackage{lineno}\linenumbers'
editor: 
  markdown: 
    wrap: sentence
execute:
  echo: false
  include: false
  eval: false
  warning: false
  message: false
---

<!--# The document is set to eval=F in the yaml. This only affect the rendering, and not the interactive session. Eval is also set explicitly for each chunk. Larger operations are cached. Remember to clean the cache occatinally and re-run. -->

@@

library(tidyverse)
library(knitr)
library(sf)
library(tmap)
library(tmaptools)
library(stars)
library(terra)
library(tidyterra)
library(ggtext)
library(cowplot)
library(units)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggmagnify)
library(ggridges)

@@

myVars <- c("7TK", "7SE", "PRTK", "PRSL", "7FA", "7GR-GI")
nts <- naturetypes_summary %>%
  rowwise() %>%
  mutate(keepers = sum(c_across(
    all_of(myVars))>0, na.rm=T)) |>
  filter(
    keepers >0,
    Ecosystem == "våtmark"
    ) |>
  pull(Nature_type)
```

@@

# Now I will copy these ADVS-values into the sf object again, keeping things in wide format
naturetypes <- naturetypes |>
  pivot_wider(
    names_from = "variable",
    values_from = "value"
  ) |>
  left_join(naturetypes_wide |> select(id, ADSV), by = "id") |>
  select(!c("7TK", "7SE", "PRSL", "PRTK"))

head(naturetypes)
```

@@

# For 7GR-GI I use this
upper2 <- 1
lower2 <- 5
threshold2 <- 2.5 # = observable effect. Value 3 indicates a shift to a new type (grunntype)

@@

```{r}
#| eval: true
# Adding scaled indicator values to the dataset
# Same code as above, but with plot=F.
naturetypes$i_ADSV <- eaTools::ea_normalise(
  data = naturetypes,
  vector = "ADSV",
  upper_reference_level = lower,
  lower_reference_level = upper,
  break_point = threshold,
  reverse = T
)

@@

# The data contains some multisurfaces 
# table(st_geometry_type(muni))
# Here is a function to make sure that multipolygons are returned
ensure_multipolygons <- function(X) {
  tmp1 <- tempfile(fileext = ".gpkg")
  tmp2 <- tempfile(fileext = ".gpkg")
  st_write(X, tmp1)
  gdalUtilities::ogr2ogr(tmp1, tmp2, f = "GPKG", nlt = "MULTIPOLYGON")
  Y <- st_read(tmp2)
  st_sf(st_drop_geometry(X), geom = st_geometry(Y))
}

@@

# and also to get the data coverage polygon.
coverage3 <- coverage |>
  st_intersection(muni3)
```

@@

```{r positionMap}
#| eval: true
# Make map to show where the three municipalities are
world <- ne_countries(scale = "medium", returnclass = "sf") |>
  st_transform(myCRS) |>
  filter(admin %in% c("Norway", "Sweden")) |>
  st_make_valid()

@@

# what is the distance between Nordre Follo and Nord-Aurdal
# First, get centroids
centroids <- muni3 |>
  st_centroid()

(km_distance <- centroids |>
  st_distance() |>
  max() |>
  set_units("km") |>
  drop_units() |>
  round())
```

@@

# I first tried to import and crop the mire data using stars, 
# but that failed (see pre 21 feb 2023).
# Trying { terra } instead

@@

# Plot to check overlap
# ggplot()+
#  geom_spatraster(data = mire_nf_terra)+
#  geom_spatvector(data = nf_vect,
#                  fill = NA)
# The cropping and masking worked.

# # I like the stars, sf and tmap combo better, so I return to stars
mire_stars_nf <- mire_terra_nf |>
  st_as_stars()
mire_stars_gr <- mire_terra_gr |>
  st_as_stars()
mire_stars_na <- mire_terra_na |>
  st_as_stars()

@@

# Make a plot to check that it has worked
(infra_dist_plot <- infraMuni3_tbl |>
  ggplot() +
  geom_bar(
    aes(
      x = infrastructureIndex,
      y = total_area_HIAs_km2,
      fill = factor(infrastructureIndex),
      colour = factor(infrastructureIndex)
    ),
    stat = "identity",
    lwd = 1.2
  ) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(4, "YlOrBr")) +
  scale_color_manual(values = RColorBrewer::brewer.pal(5, "YlOrBr")[-1]) +
  theme_minimal_hgrid() +
  labs(
    x = "Homogeneous Impact Areas",
    y = "Area (km<sup>2</sup>)"
  ) +
  theme(
    axis.title.x = element_textbox_simple(
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
      r = grid::unit(8, "pt"),
      fill = "azure1"
    ),
    axis.title.y = element_textbox_simple(
      width = NULL,
      padding = margin(4, 4, 4, 4),
      margin = margin(4, 0, 0, 0),
      linetype = 1,
      orientation = "left-rotated",
      r = grid::unit(8, "pt"),
      fill = "azure1"
    ),
    strip.background = element_blank(),
    strip.text = element_textbox(
      size = 12,
      color = "white", fill = "#5D729D", box.color = "#4A618C",
      halign = 0.5, linetype = 1, r = unit(5, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3)
    )
  ) +
  guides(fill = "none", colour = "none") +
  #scale_y_log10() +
  facet_grid(cols = vars(Municipality))
)

#ggsave(plot = infra_dist_plot,
#       "../images/infra-dist-plot.jpg")
```

```{r corrCheck}
#| eval: true
#| cache: true
# Now I want to see if the indicator values depend on the HIA is a predictable
# way to justify the stratification
corrCheck <- st_intersection(naturetypes, infra)
```

```{r HIA-validate}
#| eval: false
# A first look
corrCheck |>
  mutate(
    i_ADSV_fct = floor(round(i_ADSV * 10, 2)) / 10,
    i_alien_fct = floor(round(i_alien * 10, 2)) / 10,
    i_ditch_fct = floor(round(i_ditch * 10, 2)) / 10
  ) |>
  pivot_longer(
    cols = c(i_ADSV_fct, i_alien_fct, i_ditch_fct),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = factor(indicatorValue)
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_bw(base_size = 12) +
  guides(fill = guide_legend("Scaled indicator values")) +
  ylab("Fraction of data points") +
  xlab("HIA") +
  scale_fill_brewer(palette = "RdYlGn") +
  facet_grid(indicator ~ year)

# After this I also tried chaning the color gradient, 
# and the number of categories. I tries discrete colors and log-transformation.
# These variants are interpretted slightly differetly by the brain.
# See versions pre 27.02.2023
```

@@

@@

#tmap_save(methodsMap, "../images/studyLocations.tiff",
#           dpi= 1000,
#           units = "cm",
#           width = 18,
#           height = 10)
#tmap_save(methodsMap, "../images/studyLocations.jpg",
#           units = "cm",
#           width = 18,
#           height = 10)
#
#saveRDS(methodsMap, "../figures/studyLocation.RDS")
```

@@

# Number of mire polygons per municipality
nature3_tbl <- nature3 |>
  group_by(Municipality) |>
  summarise(n = n()) |>
  as_tibble()

# Take muni3 and add all sorts of other data to it
# using left_join.
muni_tbl <- muni3 |>
  # calculate total area
  mutate(
    area_km =
      round(
        units::drop_units(
          geom |> st_area() * 1e-6
        )
      )
  ) |>
  # make tibble for the coming join
  as_tibble() |>
  # paste inn terrestrial area
  left_join(
    terrestrial |>
      as_tibble() |>
      select(kommunenummer, t_area_km),
    keep = F
  ) |>
  # area of survey
  left_join(dk2 |> select(Municipality, dk_area_km)) |>
  mutate(dk_percent = round((dk_area_km / t_area_km) * 100)) |>
  # number of polygons
  left_join(nature3_tbl |> select(Municipality, n)) |>
  # total mire area and %
  left_join(mireArea |> select(Municipality, mirePercent, mire_km2)) |>
  # % mire inside survey coverage map
  left_join(mire_in_dk |> select(Municipality, mireInSurvey_km2)) |>
  mutate(mireInSurvery_percent = round(mireInSurvey_km2 / mire_km2 * 100, 2)) |>
  left_join(infraMuni3_summary)
```

@@

```{r forets plot}
#| eval: true
#| cache: true

# Forest plot

# #define colours for dots and bars
dotCOLS = c("grey90","grey70", "grey50", "grey40")
barCOLS <- tmaptools::get_brewer_pal("Set1", n = 4)

@@

```{r notes}
# Next I need to 
#  - copy the weighted means and SDs over to  HIA map and then over to the 
#    ecosystem delineation map polygons. 
#  - show intermittent resulting map with ggmagnify
#  - sample from these distributions (ignore missing HIAs)
```

@@

(spread_na_map <- spread_na |>
  filter(indicator == "i_ditch") |>
  mutate(Trenching = factor(round(mean, 2))) |>
  ggplot() +
  geom_sf(aes(fill=Trenching,
              color = Trenching)) +
  geom_sf(data = na,
          alpha=0) +
  scale_fill_manual(values = myCols) +
  scale_color_manual(values = myCols) +
  coord_sf(
    datum = st_crs(myCRS),
    xlim = c(494174.8 , 537114.7 ), 
    ylim = c(6737092 , 6789676)) +
  ggmagnify::geom_magnify(from = from, to = to, 
                          expand = 0,
                          shadow =T,
                          corners = 0.1) +
  theme_bw()
)
#ggsave("../images/spread-na.tiff",
#       plot = spread_na_map)
#ggsave("../images/spread-na-small.tiff",
#       plot = spread_na_map,
#       dpi=150)
#ggsave("../images/spread-na-small.jpg",
#       plot = spread_na_map)
```

@@

# Now I sample indicator values from the EDM and create a new distribution for 
# indicator values in the EAAs (the municipalities). First I sample the individual
# distributions for each mire polygon with n defined by the polygon area.
# The distribution for the polygon areas is strongly right skewed, meaning
# some polygons will contribute much more to the EAA value than others.
# I will keep this design, in-line with SEEA EA guidelines for area weighting, 
# but it's worth noting that this could be solved in other ways.
# 
combineAll <- rbind(
  spread_nf,
  spread_gr,
  spread_na
) |>
  as_tibble() |>
  drop_na() |>
  group_by(Municipality, indicator) |>
  rowwise() |>
  mutate(
    # this draws one sample pre m2 from a normal distribution:
    i_sample = list(sample(rnorm(n, mean, sd)))) |>
  select(-geometry) |>
  group_by(Municipality, indicator) |>
  reframe(i_sample = 
  # this samples randomly from i_sample, ie large polygons are more likely
  # to contibute:
            sample(
              unlist(i_sample), 
              size = 1000,
              replace = TRUE)) |>
  # truncation, since the tails from the normal distribution can go beyond 0 and 1
  mutate(i_sample = case_when(
    i_sample < 0 ~ 0,
    i_sample >1  ~ 1,
    .default = i_sample
  ))
```

```{r EEA-plot}
#| eval: true
#| cache: true
(ridgepPlot <- combineAll |>
  mutate(
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )) |>
  ggplot(
    aes(
      x = i_sample, 
      y = Municipality,
      fill = after_stat(x))
  ) +
  geom_density_ridges_gradient(
    bandwidth=.05,
    quantile_lines = TRUE, 
    quantiles = c(0.025, .5, 0.975)
  ) +
  #scale_fill_viridis_c(option = "D") +
  scale_fill_distiller(palette = "RdYlGn",
    direction = 1) +
  theme_bw(base_size = 15) +
  guides(fill = "none") +
  geom_vline(xintercept = .6,
    size=1.2,
    lty=2) +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
  labs(y = "", x = "Indicator values")+
  xlim(c(0,1)) +
  facet_wrap(.~indicator)
)

@@

#EEA_tbl |>
#  kable()
# saveRDS(EEA_tbl, "../output/EEA-table.RDS")
```
